---
title: "SDG_final 1"
author: "Huseyn Panahov"
date: "2023-05-10"
output: html_document
---

# Data wrangling 

**Revisit the observations**  

# Libraries  
```{r echo=TRUE, message=FALSE, warning=FALSE}

library(readxl) #for reading excel files
library(openxlsx)
library(dplyr) #for dataset controls/analysis  
library(ggplot2) # for visualizations 
library(plotly)
library(readr)
library(tidyr)
library(lavaan)
library(semPlot)
library(corrplot)

```

## SDG Data set  
```{r echo=TRUE, message=FALSE, warning=FALSE}

# Uploading the SDG data 
file_path <- "SDR-2022-database.xlsx"

sheet8 <- "Backdated SDG Index"

sdg_index <- read_excel(file_path, sheet = sheet8)

tail(sdg_index)
```

**Filter the data for 2021**  

```{r echo=TRUE, message=FALSE, warning=FALSE}

sdg_index_2021 <- sdg_index %>%
  filter(Year == 2021)

head(sdg_index_2021)

``` 


**Filter the data for necessary variables**  

```{r echo=TRUE, message=FALSE, warning=FALSE}
# Create a new data frame with the selected variables from sdg_index_2021
sdg_data <- sdg_index_2021 %>%
  select(Country, Region, `Income Group`, `SDG Index Score`, `Goal 1 Score`, `Goal 2 Score`, `Goal 5 Score`, `Goal 13 Score`, n_sdg6_water, n_sdg6_sanita, n_sdg7_elecac, n_sdg8_accounts, n_sdg9_intuse, n_sdg9_mobuse, n_sdg9_lpi, n_sdg11_transport, n_sdg17_statperf)

head(sdg_data)
``` 

**Optimizing the variable titles**  

```{r echo=TRUE, message=FALSE, warning=FALSE}
sdg_data <- sdg_data %>%
  rename(
    sdg_ind = 'SDG Index Score',  
    sdg1 = 'Goal 1 Score',
    sdg2 = 'Goal 2 Score', 
    sdg5 = 'Goal 5 Score', 
    sdg13 = 'Goal 13 Score',
    sdg6_water = 'n_sdg6_water', 
    sdg6_sanit = 'n_sdg6_sanita',  
    sdg7_electr = 'n_sdg7_elecac',
    sdg8_bank = 'n_sdg8_accounts',
    sdg9_internet = 'n_sdg9_intuse',
    sdg9_mobile = 'n_sdg9_mobuse',
    sdg9_lpi = 'n_sdg9_lpi', 
    sdg11_transport = 'n_sdg11_transport',
    sdg17_statperf = 'n_sdg17_statperf'
    
  )

head(sdg_data) #checking the new names 

``` 

## Merging the HCI Index  

```{r echo=TRUE, message=FALSE, warning=FALSE}
## Uploading data sets 
file_path <- "hci_data_wbg_2020.xlsx"

sheet3 <- "HCI_2020_MaleFemale"

hci_data <- read_excel(file_path, sheet = sheet3)

head(hci_data)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
## The HCI data set
hci_index <- hci_data %>%
  select('Country Name', 'HUMAN CAPITAL INDEX 2020')

hci_index <- hci_index %>%
  rename(
    Country = 'Country Name',  
    hci_ind = 'HUMAN CAPITAL INDEX 2020',
)

head(hci_index)
```

**Merging the two data frames**   

```{r echo=TRUE, message=FALSE, warning=FALSE}

# based on the "Country" column
merged_data <- merge(sdg_data, hci_index, by = "Country", all = TRUE)

tail(merged_data)
summary(merged_data)
``` 

**Removing the rows with missing sdg_ind and hci_ind variables**   

```{r echo=TRUE, message=FALSE, warning=FALSE}

new_data <- na.omit(merged_data, cols = c("sdg_ind", "hci_ind"))
summary(new_data)  

``` 


# Developing Infrastructure index     

**Kaiser-Meyer-Olkin (KMO) Test**  

The Kaiser-Meyer-Olkin (KMO) measure of sampling adequacy is a statistic that indicates the proportion of variance in the variables. The KMO values range from 0 to 1, with higher values indicating a better fit for factor analysis. The individual KMO values for each variable tell us how well each variable fits with all the others. Variables with a KMO less than 0.5 might not be suited for factor analysis, as they do not correlate well with the other variables.   
  
As we see from the below output the MSA values of all my variables are 0.8 or above, which brings the overall MSA score to 0.87. This is a positive sign, so I proceed to constructing the Infrastructure index as a latent variable.  

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(psych)

# KMO
KMO(r = cor(new_data[,c('sdg6_water', 'sdg6_sanit', 'sdg7_electr', 'sdg8_bank', 'sdg9_internet', 'sdg11_transport')])) 


``` 
  
**Side note on eigen-values**   
I checked the eigen values and based on the Kaiser-Guttman Rule, technically I could drop all the variables besides sdg6_water. It is the only with an Eigen value higher than 1. However, that is not consistent with my theoretical approach, so I proceed with my analysis. 

```{r echo=TRUE, message=FALSE, warning=FALSE}

# Eigen Values
new_data_eigen <- eigen(cor(new_data[,c('sdg6_water', 'sdg6_sanit', 'sdg7_electr', 'sdg8_bank', 'sdg9_internet', 'sdg11_transport')]))
new_data_eigen$values

``` 



**First model**  

The model above does not perform well. The CFI and TLI indicators are below the threshold, while RMSE value is very high.   

```{r echo=TRUE, message=FALSE, warning=FALSE}
infra1_model<-'#Latent Variables
infra1_index =~ sdg6_water + sdg6_sanit + sdg7_electr + sdg8_bank + sdg9_internet + sdg11_transport'

#CFA Model Estimation and Fit Measures
infra1_fit<-sem(model=infra1_model, data = new_data)
summary(infra1_fit, fit.measures=TRUE, standardized=TRUE) 

``` 
  
**Estimating the modification indices:**  

```{r echo=TRUE, message=FALSE, warning=FALSE}

infra1_fit <- sem(model=infra1_model, data = new_data)

modindices <- lavTestLRT(infra1_fit, add=TRUE)


# to calculate and sort modification indices
modindices <- modindices(infra1_fit)
sorted_modindices <- modindices[order(-modindices$mi), ]
print(sorted_modindices)


``` 

**New model with added paths:**    

```{r echo=TRUE, message=FALSE, warning=FALSE}
  
infra2_model <- '#Latent Variables
  infrastructure_index =~ sdg6_water + sdg6_sanit + sdg7_electr + sdg8_bank + sdg9_internet + sdg11_transport
  # Direct Paths
  sdg6_sanit ~ sdg6_water
  sdg9_internet ~ sdg8_bank'

# re-run the SEM analysis with this new model:
infra2_fit <- sem(model=infra2_model, data = new_data) 
summary(infra2_fit) 

``` 

**Checking the fit indices of the new model:**   

```{r echo=TRUE, message=FALSE, warning=FALSE}

summary(infra2_fit, fit.measures=TRUE)

``` 


  
**Observations:**   

This model performs much better and over it is an acceptable fit. We can see that the Comparative Fit Index (CFI) and Tucker-Lewis Index (TLI) values are above 0.99, which a very positive indicator. The Root Mean Square Error of Approximation (RMSEA) is 0.056, which is not ideal but an acceptable fit.

**Saving the new variable**   

```{r echo=TRUE, message=FALSE, warning=FALSE}

# Compute the latent variable scores
infrastr_ind_scores <- predict(infra2_fit)

# Add the latent variable scores to the original data frame 
new_data$infrastr_ind <- infrastr_ind_scores


``` 

**Rescaling the new variable: **   
The current infrastr_ind is on an unconventional scale. I use the min-max normalization technique to transform it into a new scale from 0 to 1. This method scales the values by subtracting the minimum value and dividing by the range of the original values (i.e., the difference between the maximum and minimum values). 

```{r echo=TRUE, message=FALSE, warning=FALSE}

infrastr_ind <- new_data$infrastr_ind

# Calculate the minimum and maximum values of the original variable
min_value <- min(infrastr_ind)
max_value <- max(infrastr_ind)

# Rescale the variable to the new range (0 to 1)
rescaled_var <- (infrastr_ind - min_value) / (max_value - min_value)

``` 
  
Add the new index as a new variable into my data frame: infr_ind. Saving the new data framework into a new MS Excel file.  

```{r echo=TRUE, message=FALSE, warning=FALSE}
new_data$infr_ind <- rescaled_var 

write.xlsx(new_data, "sdg_hci_data3.0.xlsx")
``` 


```{r,echo=FALSE, message=FALSE, warning=FALSE}

``` 



```{r,echo=FALSE, message=FALSE, warning=FALSE}


``` 


# 

```{r,echo=FALSE, message=FALSE, warning=FALSE}

``` 


# 

```{r,echo=FALSE, message=FALSE, warning=FALSE}

``` 
